Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-16 ieee_paper.tex
section 3 complete?

Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios in
the best and the worst case.

@@

  \caption{Ratio of flops to bytes: in the best case in
    \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
    \ref{fig:ratio_worst}; and in the worst case neglecting that the
    memory transfers are by batch of size $cl$ in
    \ref{fig:ratio_worst_nocache}}
  \label{fig:ratio_bytes_flops}
\end{figure*}

\begin{figure*}
  \centering 
  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
%
  \subfigure[Relative performance to base case using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{Estimation of the reachable performance one can
      achieve using a device with a memory to computational units of
      150GB/s when varying the number of vectors and matrices}
  \label{fig:perf_predict}
\end{figure*}

\todo{conclude on the evaluation using the figure/table. Note that we
  base prediction on 150GB/s which is not the peak of saule12. Note
  that if you are in best in v1m1 you might go to bad v4m4, but if you
  are in bad v1m1, you will essentially stay there.}

>>>>

Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
for single precision computation in the best and the worst case on a
classical cache based architecture ($cl = 64$) and assuming $cl=1$. We
can easily see the potential improvement in the computational
intensity when the number of matrices of vectors increase. There is a
significant difference between the best and the worst case: there is a
$8$ fold difference in the 1 matrix 1 vector case but that gap closes
with the increase in the number of vectors and matrices to a 4 fold
difference in the 4 matrices and 4 vectors case and 1.6 fold in the 16
matrices and 16 vectors. The ratios computed with $cl = 1$ are mostly
similar to the the one with $cl=64$ when the number of vectors is
large. But the differences are important when the number of vector is
low: this highlight that accessing the memory per batch is the main
problem faced when performing a standard SpMV computation.

@@

  \caption{Ratio of flops to bytes in single precision: in the best case in
    \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
    \ref{fig:ratio_worst}; and in the worst case neglecting that the
    memory transfers are by batch (equivalent to $cl = 1$) in
    \ref{fig:ratio_worst_nocache}.}
  \label{fig:ratio_bytes_flops}
\end{figure*}

Figure~\ref{fig:ratio_bytes_flops} shows how these ratios translate
into actual performance. This figure provides projected best and worst
Gflop/s achievable assuming the computation is memory bound and the
architecture reaches 150Gflop/s. Notice that \cite{Saule13-ARXIV}
showed a higher peak bandwidth, but we will show in
Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
might achieve in these kernels. In single precision, the worse that
one could achieve by using 4 vectors and 4 matrices is much higher
than the best achievable using a classical SpMV computation. The best
performance reachable is 210Gflop/s: almost 6 times higher than the
peak of the classical SpMV case.

\begin{figure*}
  \centering 
  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
%
  \subfigure[Relative performance to using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{Estimation of the reachable performance one can achieve
    using a device with a memory to computational units of 150GB/s
    when varying the number of vectors and matrices. \todo{we could
      remove the speedup plot if we need space.}  \todo{can the legend
      read "Single Precision (best)"} \todo{there is a missing
      parenthesis in the title; alternatively remove title.}}
  \label{fig:perf_predict}
\end{figure*}
